---
title: "assignment_6"
format: gfm
---

```{r}
library(tidyverse)
library(knitr)
```

## **Exercise 1. Tibble and Data Import**

Import the data frames listed below into R and [parse](https://r4ds.had.co.nz/data-import.html#parsing-a-vector) the columns appropriately when needed. Watch out for the formatting oddities of each dataset. Print the results directly, **without** using `kable()`.

**You only need to finish any three out of the five questions in this exercise in order to get credit.**

#### **1.1 Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results. \[We didn’t have time to cover this in class, but look up how these functions work [here](https://r4ds.had.co.nz/tibbles.html#creating-tibbles)\]**

`tribble()`:

```{r}
tibble(a=1:2, b=c(2.1,3.2), c=c("apple", "orange"))
```

`tibble()`:

```{r}
tribble(~a, ~b, ~c, 1, 2.1, "apple", 2, 3.2, "orange")
```

#### **1.2 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.**

```{r}
#| message: false
#| warning: false
#| paged-print: false
dataset2 <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt", col_names = c("Name", "Weight", "Price"))

dataset2
```

#### **1.3 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt` into R. Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.**

```{r}
#| message: false
#| warning: false
#| paged-print: false
dataset3 <- read_delim ("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt", delim=";", skip=3, col_names = c("Name", "Weight", "Price"), na = c("Not Available", "?")) |>
  mutate (Name = str_remove_all(Name, "/"))

dataset3
```
