---
title: "lab 9"
format: html
---

```{r}
library(tidyverse)
```

#### **1.1 Write a function, `dna_or_rna(sequence)`, that determines if a sequence of base pairs is DNA, RNA, or if it is not possible to tell given the sequence provided.**

-   Since all the function will know about the material is the sequence, the only way to tell the difference between DNA and RNA is that RNA has the base Uracil (`"u"`) instead of the base Thymine (`"t"`).

-   Have the function return one of three outputs: “DNA”, “RNA”, or “unknown”. Then run the following three lines of code:

`dna_or_rna("attggc")`

`dna_or_rna("gccaau")`

`dna_or_rna("ccagac")`

`dna_or_rna("tgcacug")`

Hint: the `str_split` function might be helpful.

```{r} function(sequence){}
     base <- strsplit(sequence, pattern = "") |>
       unlist() |>
       unique()
     if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
       return("DNA")
     } else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
       return("RNA")
     } else {
       return("unknown")
     }
}

```

```         
## function(sequence){
##   bases <- str_split(sequence, pattern = "") %>%
##     unlist() %>%
##     unique()
##   if (all(bases %in% c("a", "t", "g", "c")) & "t" %in% bases){
##     return("DNA")
##   } else if (all(bases %in% c("a", "u", "g", "c")) & "u" %in% bases){
##     return("RNA")
##   } else {
##     return("unknown")
##   }
## }
## <bytecode: 0x1359d40a0>
```
